---
title: "Analyze ESAS questionnaire for draft report"
author: "Lien Reyserhove"
output: html_document
---

# General settings

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

Load libraries

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(magrittr)
library(here)
library(knitr)
library(formattable)
library(sf)
library(leaflet)
library(htmltools)
library(htmlwidgets)
library(webshot)
```

# Import data

```{r import data}
input_data <- read.delim(file = here::here("data", "intermediate", "input_data_cleaned.csv"), sep = ",")
```

Import spatial information dataset for map:

```{r import geojson}
global_countries_spatial <- st_read(here::here("./data/input/CNTR_RG_60M_2020_4326.geojson"))
```

# Paragraph 1: General project information

## Figure 1: Number of datasets per country

Generate `datasets_per_country`

```{r}
datasets_per_country <- 
  input_data %>% 
    select(country_clean) %>% 
    group_by(country_clean) %>% 
    summarize(n_datasets = n())
datasets_per_country
```

filter spatial information on countries of interest

```{r}
focal_countries_spatial <- 
  global_countries_spatial %>% 
    filter(NAME_ENGL %in% input_data$country_clean)
```


Join information with `countries_spatial`

```{r}
focal_countries_spatial %<>% 
  left_join(datasets_per_country,
            by = c("NAME_ENGL" = "country_clean"))
```

Plot information:

```{r}
bins <- c(1, 2, 3, 4)
pal <- colorNumeric("YlOrRd", domain = focal_countries_spatial$n_datasets)

figure_1 <-
  leaflet(focal_countries_spatial) %>%
  setView(lng = 9,
          lat = 51,
          zoom = 3.75) %>%
  addTiles() %>%
  addPolygons(
    fillColor = ~pal(n_datasets),
    weight = 2,
    opacity = 1,
    color = "grey",
    dashArray = "3",
    fillOpacity = 1,
    highlight = highlightOptions(
      weight = 5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE),
    label = ~htmlEscape(n_datasets)) %>%
  addLegend("bottomright",
            pal = pal, values = ~n_datasets,
            title = "number of SAS datasets",
            opacity = 1)
figure_1
```

Save html via Export -> Save as web page or via saveWidget:

```{r}
saveWidget(figure_1,
           "figure_1.html",
           selfcontained = TRUE)
```

```{r}
webshot("figure_1.html",
        file = "../data/processed/figure_1.png",
        cliprect = "viewport")
```

## Table 1: General project information

This table discusses the following dataset properties:

- Dataset ID
- Country
- Organization (acronym)
- Programme name and/or acronym
- active (yes/no)
- esas database (yes/no)

```{r}
table_1 <- 
  input_data %>%
      select(dataset_id, country_clean, organization_acronym, name_acronym_clean, esas_db_yn, active_yn) %>% 
  rename(`Dataset ID` = dataset_id) %>%
  rename(`Country` = country_clean,
         `Organization (acronym)` = organization_acronym,
         `Programme name and/or acronym` = name_acronym_clean,
         `ESAS database (y/n)` = esas_db_yn,
         `Active (y/n)` = active_yn) %>% 
      arrange(`Country`) %>% 
      kbl() %>% 
      kable_classic(full_width = T, html_font = "Arial") %>% 
      kable_styling(bootstrap_options = c("striped", "hover")) 
table_1
```

```{r}
 as_image(table_1, file = "../data/processed/table_1.png")
```

## Text

### Integration in ESAs database?

Summarize `esas_db_yn`:

```{r}
input_data %>% 
  select(esas_db_yn) %>% 
  group_by_all() %>% 
  summarize(records = n(),
            percentage = round(n()/nrow(input_data)*100))
```

### Programme active y/n?

Summarize `active_yn`:

```{r}
input_data %>% 
  select(active_yn) %>% 
  group_by_all() %>% 
  summarize(records = n(),
            percentage = round(n()/nrow(input_data)*100))
```

# Paragraph 2: Data content

# Paragraph 3: Constraints for sharing data

### Table 3

This table discusses the following dataset properties:

- Sharing interest (Y/n)
- Financial constraints (y/n)
- Legal constraints (y/n)

### Sharing interest

Screent content of `sharing_interest_yn`:

```{r}
input_data %>%
  select(sharing_interest_yn) %>% 
  group_by(sharing_interest_yn) %>% 
  summarize(records = n())
```

No cleaning required

### Financial constraints

Screent content of `financial_constraints_yn`:

```{r}
input_data %>%
  select(financial_constraints_yn) %>% 
  group_by(financial_constraints_yn) %>% 
  summarize(records = n())
```

In table 3, it is better to reformulate to `financial_condition` which should be `OK`(thus `financial_constraints`=`No`), `NOK`(thus `financial_constraints`=`Yes`) or `Unknown`

```{r}
input_data %<>% 
  mutate(financial_condition = case_when(
    financial_constraints_yn == "Yes" ~ "NOK",
    financial_constraints_yn == "No" ~ "OK",
    financial_constraints_yn == "Unknown" ~ "-"))
```

### Legal constraints

In table 3, it is better to reformulate to `legal_condition` which should be `OK`(thus `legal_constraints`=`No`), `NOK`(thus `legal_constraints`=`Yes`) or `Unknown`

```{r}
input_data %<>% 
  mutate(legal_condition = case_when(
    legal_constraints_yn == "Yes" ~ "NOK",
    legal_constraints_yn == "No" ~ "OK",
    legal_constraints_yn == "Unknown" ~ "-"))
```

### Summary

For this table we use the `formattable` package. FOr this, we need to define the formatters for `y/n` and `OK/NOK` first:

Define formatters:

```{r}
yn_formatter <-
  formatter("span",
            style = 
              x ~ style(color = ifelse(x == "Yes", "green", 
                                       ifelse(x == "No", "red", "grey"))),
              x ~ icontext(ifelse( x == "Yes", "ok", 
                                   ifelse(x == "No", "remove", "minus")),
                           ifelse(x == "Yes", "Yes",
                                  ifelse(x == "No", "No", ""))),
            "font_size" = 40)

OK_NOK_formatter <-
  formatter("span",
            style = 
              x ~ style(color = ifelse(x == "OK", "green", 
                                       ifelse(x == "NOK", "red", "grey"))),
              x ~ icontext(ifelse( x == "OK", "ok", 
                                   ifelse(x == "NO", "remove", "minus")),
                           ifelse(x == "OK", "OK",
                                  ifelse(x == "NOK", "NOK", ""))),
            "font_size" = 40)
```

Generate table 3:

```{r}
table_3 <- 
  input_data %>% 
    select(dataset_id, sharing_interest_yn, financial_condition, legal_condition) %>%
    rename(`Dataset ID` = dataset_id,
           `Sharing interest (y/n)` = sharing_interest_yn,
           `financial condition (OK/NOK))` = financial_condition,
           `legal condition (OK/NOK)` = legal_condition) %>% 
    arrange(`Dataset ID`) 
```


```{r}
formattable(table_3,
            align = c("l","c","c","c"),
            list(
              `Sharing interest (y/n)` = yn_formatter,
              `financial condition (OK/NOK))` = OK_NOK_formatter,
              `legal condition (OK/NOK)` = OK_NOK_formatter)) %>% 
  as_image(file = "../data/processed/table_3.png")
```

