---
title: "First inspection ESAS questionnaire"
author: "Lien Reyserhove"
date: "25 augustus 2020"
output: html_document
---


# General settings

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

Load libraries

```{r, warning=FALSE, message=FALSE}
library(googlesheets4)
library(tidyverse)
library(janitor)
library(magrittr)
library(here)
library(knitr)
library(kableExtra)
```

```{r}
gs4_deauth()
```

# Import data

Import google sheet

```{r, message=FALSE}
input_data <- read_sheet("https://docs.google.com/spreadsheets/d/1uR4z8UYCje7CfqYDezbBlBIiSsvyyalfSPwWZEBfcIQ/edit#gid=1639204022", col_types = "c")
```

Save raw dataset to GitHub:

```{r}
write.csv(input_data, here::here("data", "input", "input_data.csv"), row.names = FALSE)
```

# Data cleaning

## Clean column names

Actions:

- Clean columns names

```{r, echo=FALSE, message=FALSE}
input_data %<>% clean_names() 
```

- Rename headers

```{r, echo=FALSE}
input_data %<>% rename(
  "sas_data_yn"               = "has_your_organization_monitored_seabirds_at_sea_past_or_present",
  "name_and_acronym"          = "name_and_or_acronym_of_the_monitoring_programme_if_not_specified_please_put_none",
  "short_description"         = "short_description_or_url_of_the_purpose",
  "programm_lead"             = "organization_s_leading_or_commissioning_the_monitoring_programme",
  "country"                   = "country_where_this_organization_is_based",
  "role"                      = "what_is_your_role_in_this_monitoring_programme",
  "active_yn"                 = "is_this_monitoring_programme_currently_active",
  "protocol_yn"               = "data_are_collected_according_to_esas_protocol",
  "nr_occurrences"                = "number_of_species_occurrence_records_i_e_species_key_records",
  "data_storage"              = "data_are_stored_as",
  "data_import"               = "data_are_imported_into_above_system",
  "language"                  = "language_use_in_data_e_g_english_dutch",
  "esas_db_yn"                = "data_are_included_in_esas_database",
  "data_sample__yn"           = "can_you_provide_a_sample_of_the_data_to_help_us_assess_data_compatibility",
  "other_sas_data"            = "are_you_aware_of_other_seabird_at_sea_monitoring_programmes_in_your_country_please_provide_contact_information_if_possible",
  "sharing_interest_yn"       = "is_there_any_interest_for_sharing_data_with_esas",
  "financial_constraints_yn" = "are_there_any_financial_constraints_for_sharing_data_with_esas",
  "legal_constraints_yn"    = "are_there_any_intellectual_ownership_constraints_for_sharing_data_with_esas",
  "db_model_yn"               = "are_you_aware_of_the_current_esas_database_model",
  "shortcomings_esas_model"   = "if_your_answer_to_the_previous_question_was_yes_can_you_specify_potential_shortcomings_of_the_current_esas_database_model_for_integrating_your_data",
  "expected_costs"            = "if_your_answer_to_the_previous_question_was_yes_can_you_give_an_estimation_of_the_expected_cost_to_incorporate_the_data_in_the_esas_database",
  "ship_based_line_transect" = "if_your_monitoring_programme_includes_ship_based_observations_please_specify_the_specific_method_line_transect_methodology",
  "ship_based_strip_transect" = "if_your_monitoring_programme_includes_ship_based_observations_please_specify_the_specific_method_strip_transect_methodology",
  "ship_based_flying_birds"   = "if_your_monitoring_programme_includes_ship_based_observations_please_specify_the_specific_method_snapshot_count_for_flying_birds",
  "ship_based_distance"       = "if_your_monitoring_programme_includes_ship_based_observations_please_specify_the_specific_method_distance_estimation",
  "aerial_digital_survey"     = "if_your_monitoring_programme_includes_aerial_observations_please_specify_the_specific_method_digital_survey",
  "aerial_observer_based"     = "if_your_monitoring_programme_includes_aerial_observations_please_specify_the_specific_method_observer_based_survey",
  "aerial_line_transect"      = "if_your_monitoring_programme_includes_aerial_observations_please_specify_the_specific_method_line_transects",  
  "aerial_strip_transect"     = "if_your_monitoring_programme_includes_aerial_observations_please_specify_the_specific_method_strip_transects",  
  "distance_estimation"       = "if_your_monitoring_programme_includes_aerial_observations_please_specify_the_specific_method_distance_estimation",
  "sharing_conditions"        = "if_your_answer_to_the_previous_question_was_yes_can_you_specify_what_those_constraints_are_and_or_under_what_conditions_data_could_be_shared",  
  "information_about"         = "does_your_monitoring_programme_include_information_about")
```

Remove responses for which  `sas_data_yn` = `No` (no SAS data available):

```{r remove non sas data}
input_data %<>% filter(sas_data_yn != "No")
```

## Clean country information

Country information is key in the discussion of the report. We need to clean this information.

Show country names:

```{r}
input_data %>% distinct(country)
```

For one dataset, the field `country` included two countries: `Norway and Russia`. As this information is linked to the Norwegian Institute for Nature Research, we consider the origin of this dataset to be Norwegian. Additionally, we clean `Eesti`, provide the full name for `UK` and use title case for all country names:

```{r}
input_data <-
  input_data %>% 
    mutate(country_clean = 
      recode(country, 
      "Norway and Russia" = "Norway",
      "Eesti" = "Estonia",
      "UK" = "United Kingdom",
      "Russia" = "Russian Federation")) %>% 
    mutate(country_clean = str_to_title(country_clean))
```

Compare original and cleaned country names:

```{r}
input_data %>% 
  select(country, country_clean) %>% 
  group_by_all() %>% 
  summarize(n = n())
```

## Organizations 

Clean `organizaton` to `organization_acronym_clean`

```{r}
input_data %<>% mutate(organization_acronym = recode(organization,
  "INBO" = "Research Institute for Nature and Forest (INBO)",
  "RBINS" = "Royal Belgian Institute for Natural Sciences (RBINS)",
  "Estonian University of Life Sciences" = "Estonian University of Life Sciences (EMU)", #source: https://www.emu.ee/en/
  "BirdLife Estonia" = "BirdLife Estonia (EOS)", #source:https://www.birdlife.org/europe-and-central-asia/partners/estonia-%E2%80%93-estonian-ornithological-society-eos
  "BSH Federal Maitime and Hydrographic Agency" = "Federal Maritime and Hydrographic Agency (BSH)", #source = https://www.bsh.de/EN/Home/home_node.html
  "FTZ, Kiel University" = "Kiel University, Research and Technology Centre Westcoast (FTZ)", # source = https://www.ftz.uni-kiel.de/en/about-us
  "National Parks and Wildlife Service" = "National Parks and Wildlife Service (NPWS)", #source = https://www.npws.ie/
  "Latvian Ornithological society" = "Latvian Ornithological society (LOB)", #source = https://www.birdlife.org/europe-and-central-asia/partners/latvia-%E2%80%93-latvian-ornithological-society-lob
  "Nature Research Centre" = "Nature Research Centre (NRC)", #source = http://www.gamtostyrimai.lt/en/
  "Bureau Waardenburg" = "Bureau Waardenburg (BuWa)",
  "Royal Netherlands Institute for Sea Research" = "Royal Netherlands Institute for Sea Research (NIOZ)", #source = https://www.nioz.nl/en,
  "Norwegian Institute for Nature Research" = "Norwegian Institute for Nature Research (NINA)", #source = https://www.nina.no/english
  "Econ Ecological Consultancy" =  "Ecological Consultancy (ECON)", #source = https://www.econ-ecology.com/
  "Chief Inspector of Environmental Protection" = "Chief Inspector of Environmental Protection (GIOS)", #source = http://www.gios.gov.pl/en/
  "Department of Vertebrate Ecology & Zoology, University of Gdansk, Poland" = "Department of Vertebrate Ecology & Zoology, University of Gdansk, Poland (KEZK)", #source = https://kezk.bio.ug.edu.pl/index.php?lang=pl
  "SPEA" = "Society for the Study of Birds (SPEA)", #source = https://www.spea.pt/en/about-us/spea/
  "Saint-Petersburg Research Center of Russian Academy of sciences" = "Scientific Research Center for Ecological Safety (SRCES RAS)", #source = http://www.istc.int/en/institute/9789
  "AZTI" = "AZTI Tecnalia (AZTI)", #source = https://www.euromarinenetwork.eu/membership/organisations/azti-tecnalia
  "HiDef Aerial Surveying Limited" = "HiDef Aerial Surveying Limited (HiDef)", #source = https://hidef.bioconsult-sh.de/
  "JNCC" = "Joint Nature Conservation Committe (JNCC)", #source =
  "Econ Ecological Consultancy Ltd" = "Ecological Consultancy (ECON)", #source = https://www.econ-ecology.com/,
  "Hellenic Ornithological Society" = "Hellenic Ornithological Society (HOS)", #
  "Hellenic Ornithological Society / BirdLife Greece" = "Hellenic Ornithological Society (HOS) / BirdLife Greece" #https://www.ornithologiki.gr/index.php?loc=en
  )) 
```

Total number of organizations holding SAS data:

```{r}
input_data %>% 
  distinct(organization_acronym) %>% 
  summarize(records = n())
```

## Name and acronym

Clean `name_and_acronym` to `name_acronym_clean`

```{r}
input_data %<>% mutate(programme_acronym = recode(name_and_acronym,
  "German Seabirds at Sea programme" = "German Seabirds at Sea programme (SAS)", 
  "Multiple post-consent and SPA monitoring projects" = "Multiple post-consent and Special Protection Area (SPA) monitoring projects",
  "SAS" = "Brittish Seabirds at Sea programme (SAS)",
  "LIFE Marine SPAs" = "LIFE Marine Special Protection Areas (SPA)",
  "MWTL zeevogels" = "Monitoring van de Waterstaatkundige Toestand des Lands (MWTL), zeevogels",
  "None" = "",
  "None. We do not do monitoring" = "",
  "Sheringham Shoal OWF" = "Sheringham Shoal Offshore Wind Farm (OWF)",
  "SEAPOP" = "Seabird populations (SEAPOP)",
  "Monitoring of Wintering Sea Birds (PL: MZPM - Monitoring Zimujacych Ptaków Morskich)" = "Monitoring of Wintering Sea Birds (MZPM)",
  "MZPM" = "Monitoring of Wintering Sea Birds (MZPM)",
  "ESAS" = "European Seabirds at Sea (ESAS)",
  "Conservation measures for Falco eleonorae in Greece (LIFE03 NAT/GR/000091" = "Conservation measures for Falco eleonorae in Greece (LIFE03 NAT/GR/000091)"
))
```

Compare `name_and_acronym` with `name_acronym_clean`

```{r}
input_data %>% 
  select(name_and_acronym, programme_acronym) %>% 
  group_by_all() %>% 
  summarize(records = n())
```



## Dataset ID

To link all dataset properties to the right dataset, we generate a dataset identifier here. This will be a combination between the country's ISO code, institution acronym (if not available we make one ourselves) and (in case of multiple datasets per country) dataset acronym.

Add dataset shortname:

```{r generate shortname}
input_data %<>% mutate(dataset_id = case_when(
  
  # Belgium
  country_clean == "Belgium" & organization_acronym == "Research Institute for Nature and Forest (INBO)" ~ "BEL-INBO",
  country_clean == "Belgium" & organization_acronym == "Royal Belgian Institute for Natural Sciences (RBINS)" ~ "BEL-RBINS",
  
  # Estionia
  country_clean == "Estonia" & organization_acronym == "BirdLife Estonia (EOS)" ~ "EST-EOS",
  country_clean == "Estonia" & organization_acronym == "Estonian University of Life Sciences (EMU)"  ~"EST-EMU",
  
  # France
  country_clean == "France" ~ "FRA",
  
  # Germany
  country_clean == "Germany" & 
    organization_acronym == "Federal Maritime and Hydrographic Agency (BSH)" 
      ~ "DEU-BSH",
  country_clean == "Germany" & 
    organization_acronym == "Kiel University, Research and Technology Centre Westcoast (FTZ)" &
    programme_acronym == "German Marine Biodiversity Monitoring" 
      ~ "DEU-FTZ-GMBM",
  country_clean == "Germany" & 
    organization_acronym == "Kiel University, Research and Technology Centre Westcoast (FTZ)" &
    programme_acronym == "German Seabirds at Sea programme (SAS)" 
      ~ "DEU-FTZ-SAS",
  country_clean == "Germany" & 
    organization_acronym == "Kiel University, Research and Technology Centre Westcoast (FTZ)" &
    programme_acronym == "Offshore Seabird Monitoring in the Schleswig-Holstein coastal sea, North Sea" 
      ~ "DEU-FTZ-OSMS",
  
  # Greece
  country_clean == "Greece" & 
    organization_acronym == "Hellenic Ornithological Society (HOS)" & 
    programme_acronym == "Actions for the Conservation of Larus audouinii in Greece (LIFE96 NAT/GR/003221)"
      ~ "GRC-HOS-LIFE96",
  
   country_clean == "Greece" & 
    organization_acronym == "Hellenic Ornithological Society (HOS) / BirdLife Greece" & 
    programme_acronym == "Conservation measures for Falco eleonorae in Greece (LIFE03 NAT/GR/000091)" 
      ~ "GRC-HOS-LIFE03",
  
  country_clean == "Greece" & 
    organization_acronym == "Hellenic Ornithological Society (HOS) / BirdLife Greece" & 
    programme_acronym == "Concrete conservation measures for the Mediterranean Shag and Audouin’s Gull in Greece, including the inventory of relevant marine IBAs (LIFE07 NAT/GR/000285)" 
      ~ "GRC-HOS-LIFE07",
  
  country_clean == "Greece" & 
    organization_acronym == "Hellenic Ornithological Society (HOS) / BirdLife Greece" & 
    programme_acronym == "Hellenic Seabird Project" 
      ~ "GRC-HOS-HSB",
  
  country_clean == "Greece" & 
   organization_acronym == "Hellenic Ornithological Society (HOS) / BirdLife Greece" & 
   programme_acronym == "EU Marine Strategy Framework Directive for Greece" 
      ~ "GRC-HOS-EU",

  # Ireland
  country_clean == "Ireland" & 
    organization_acronym == "Marine Institute" ~ "IRL-MI",
  country_clean == "Ireland" & 
    organization_acronym == "National Parks and Wildlife Service (NPWS)" ~ "IRL-NPWS",
 
  # Latvia
  country_clean == "Latvia" ~ "LVA",

  # Lithuania
  country_clean == "Lithuania" ~ "LTU",

  # Malta
 country_clean == "Malta" ~ "MLT",

  # Netherlands
  country_clean == "Netherlands" & 
   organization_acronym == "Bureau Waardenburg (BuWa)" & 
   name_and_acronym == "None" &
   platform == "Aerial" 
    ~ "NLD-BuWa-AIR",
  country_clean == "Netherlands" & 
   organization_acronym == "Bureau Waardenburg (BuWa)" & 
  name_and_acronym == "None" &
   platform == "Ship-based" 
    ~ "NLD-BuWa-SHIP",
  country_clean == "Netherlands" & 
   organization_acronym == "Royal Netherlands Institute for Sea Research (NIOZ)" 
    ~ "NLD-NIOZ", 
   country_clean == "Netherlands" & 
    organization_acronym == "Bureau Waardenburg (BuWa)" & 
    name_and_acronym == "MWTL zeevogels"
    ~ "NLD-BuWA-MWTL",
 
  # Norway
  country_clean == "Norway" & 
   organization_acronym == "Ecological Consultancy (ECON)" 
    ~ "NOR-ECON",
  country_clean == "Norway" & 
   organization_acronym == "Norwegian Institute for Nature Research (NINA)" 
    ~ "NOR-NINA",

  # Poland
  country_clean == "Poland" & 
   organization_acronym == "Chief Inspector of Environmental Protection (GIOS)" 
    ~ "POL-MZPM-1",
  country_clean == "Poland" & 
   organization_acronym == "Department of Vertebrate Ecology & Zoology, University of Gdansk, Poland"
    ~ "POL-MZPM-2",
  
  # Portugal
  country_clean == "Portugal" ~ "PRT",

  # Russian Federation
  country_clean == "Russian Federation" ~ "RUS",

  # Spain
  country_clean == "Spain" & 
   organization_acronym == "AZTI Tecnalia (AZTI)" & 
   programme_acronym == "BIOMAN" 
    ~ "ESP-AZTI-BIO",
  country_clean == "Spain" & 
   organization_acronym == "AZTI Tecnalia (AZTI)" & 
   programme_acronym == "JUVENA" 
    ~ "ESP-AZTI-JUV",
  country_clean == "Spain" & 
   organization_acronym == "Instituto Español de Oceanografía (IEO)" 
    ~ "ESP-IEO",
 
  # Sweden
  country_clean == "Sweden" ~ "SWE",

  # United Kingdom
  country_clean == "United Kingdom" & 
   organization_acronym == "Ecological Consultancy (ECON)" 
    ~ "GBR-ECON",
  country_clean == "United Kingdom" & 
   organization_acronym == "HiDef Aerial Surveying Limited (HiDef)" 
    ~ "GBR-HiDef",
  country_clean == "United Kingdom" & 
   organization_acronym == "Joint Nature Conservation Committe (JNCC)" 
    ~ "GBR-JNCC"
  ))
```

Inspect dataset shortnames per country:

```{r inspect dataset shortname}
input_data %>% 
  select(country_clean, organization_acronym, programme_acronym, dataset_id) %>% 
  group_by_all() %>% 
  summarize(records = n()) %>% 
  kbl()
```

# Export cleaned dataset

Resturecure and Select columns relevant for further discussion (to reduce size of the dataset)

```{r}
input_data %<>% 
  select(
    dataset_id,
    country_clean,
    organization_acronym,
    programme_acronym,
    role,
    email_address,
    short_description,
    programm_lead,
    active_yn,
    start_year,
    end_year,
    nr_occurrences,
    taxonomic_scope,
    geographical_scope,
    time_of_year,
    sampling_interval,
    platform,
    data_storage,
    data_import,
    esas_db_yn,
    remarks,
    other_sas_data,
    sharing_interest_yn,
    financial_constraints_yn,
    legal_constraints_yn,
    expected_costs,
    sharing_conditions,
    information_about,
    other_relevant_remarks,
    ship_based_line_transect,
    ship_based_flying_birds,
    ship_based_distance,
    aerial_digital_survey,
    aerial_observer_based,
    aerial_line_transect,
    aerial_strip_transect,
    distance_estimation,
    ship_based_strip_transect
  )
```

Export cleaned dataset: 

```{r}
write.csv(input_data, here::here("data", "intermediate", "input_data_cleaned.csv"), row.names = FALSE)
```
